\svnInfo $Id$

\subsection{Applicable documents}
\label{sec:applicable documents}

Table \ref{tab:applicable list} lists all the LOFAR ICDs.  Most of the ICDs are for the various LOFAR data types, while ICD numbers 002 and 005 are general and applicable to all the data-format-oriented ICDs.  Please note that the data and header information is written in Little-endian format within the HDF5 files.

\begin{center}
  %% Table head
  \tablefirsthead{
    \hline
    \textsc{Reference} & \textsc{Title} & \textsc{Description} \\
    \hline
  }
  \tablehead{
    \multicolumn{3}{r}{\small \textbf{Applicable documents}
      \textsl{continued from previous page}} \\
    \hline
    \textsc{Reference} & \textsc{Title} & \textsc{Description} \\
    \hline
  }
  %% Table tail
  \tabletail{
    \hline
    \multicolumn{3}{r}{\small\sl continued on next page} \\
  }
  \tablelasttail{\hline}
  %% Table contents
  \bottomcaption[List of all LOFAR Interface Control Documents (ICDs)]{
    List of all the LOFAR Interface Control Documents.  ICDs 001, 003, 004, 006, 007 and 008 describe different LOFAR data
    formats, while ICDs 002 and 005 are general and apply to the other ICDs.
  \label{tab:applicable list}}
  \begin{supertabular}{lp{5.8cm}p{7cm}}
    \shrinkheight{-4.5em}
    \texttt{ICD-001} \cite{lofar.icd.001} & TBB Time-Series Data & Digitized
    voltage output, as received by the individual LOFAR dipoles. \\
    \texttt{ICD-002} \cite{lofar.icd.002} & Representations of World
    Coordinates & Definition of how to represent and store metadata
    that serve to locate a measurement in some multidimensional parameter space. \\
    \texttt{ICD-003} \cite{lofar.icd.003} & Beam-Formed Data & Hosting
    structure for LOFAR Beam-Formed data. \\
    \texttt{ICD-004} \cite{lofar.icd.004} & Radio Sky Image Cubes &
    Primary data product of the imaging pipeline. \\
    \texttt{ICD-005} \cite{lofar.icd.005} & File Naming Conventions &
    Conventions for the naming scheme applied to LOFAR standard data products. \\
    \texttt{ICD-006} \cite{lofar.icd.006} & Dynamic Spectrum Data &
    Hosting structure for dynamic spectrum data, i.e. intensity as
    function of time and frequency. \\
    \texttt{ICD-007} \cite{lofar.icd.007} & Visibility Data & Hosting structure for
    LOFAR UV Visibility data, primary output of interferometer operations. \\
    \texttt{ICD-008} \cite{lofar.icd.008} & RM Synthesis Cubes &
    Hosting structure for LOFAR Rotation Measure Synthesis Cubes
    output data. \\
    \hline
  \end{supertabular}
\end{center}
